SEER Medicaid

### 2023-12-04 Manuscript Analysis code 0 to 19.Rmd
adding revision changes

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

@@

seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("In situ/localized", "Regional", "Distant"))

@@

# test code
```{r}

# create categories for AYA codes
class(seer0to39$AYA_site_recode_WHO_2008) # numeric

seer0to39 <- seer0to39 %>%
  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
                                 AYA_site_recode_WHO_2008 %in% c(55,56) ~ 9,
                                 AYA_site_recode_WHO_2008 == 99 ~ 10))

seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
  "1. Leukemias",
  "2. Lymphomas",
  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
  "4. Osseous & Chondromatous Neoplasms",
  "5. Soft Tissue Sarcomas",
  "6. Germ Cell and Trophoblastic Neoplasms",
  "7. Melanoma and Skin Carcinomas",
  "8. Carcinomas",
  "9. Miscellaneous specified neoplasms, NOS",
  "10. Unspecified Malignant Neoplasms",
  "Unclassified"))
table(seer0to39$can_typeAYA, useNA = "always")

# categorize brain tumors according to AYA definition

seer0to39 <- seer0to39 %>%
  mutate(brain = case_when (can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 %in% c(7,8,9) ~ 0, # 3.1 Astrocytoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(10) ~ 1, # 3.2 Other glioma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(11) ~ 2, # 3.3 Ependymoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(12,13) ~ 3, # 3.4 Medulloblastoma and other PNET
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(14) ~ 4, # 3.5 Other specified intracranial and instraspinal neoplasms
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(15,16) ~ 5)) # 3.6 Unspecified intracranial and intraspinal neoplasms
seer0to39$brain
table(seer0to39$brain, useNA = "always")

seer0to39$brain <- factor(seer0to39$brain, levels = c(0:5), labels = c(
  "3.1 Astrocytoma",
  "3.2 Other glioma",
  "3.3 Ependymoma",
  "3.4 Medulloblastoma and other PNET",
  "3.5 Other specified intracranial and instraspinal neoplasms",
  "3.6 Unspecified intracranial and intraspinal neoplasms"))
table(seer0to39$brain, useNA = "always")

df <- seer0to39 %>%
  filter(can_type == "III. CNS and miscellaneous intracranial and intraspinal neoplasms"|can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)") %>%
  group_by(can_type, can_typeAYA)%>%
  tally()
```

@@

# Exclude those diagnosed before 2006 (a handful)
```{r}
seer0to19 <- seer0to19 %>%
  filter(Year_of_diagnosis >=2006) # 32,380
```

@@

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) # Other
table(seer0to19$mcont66, useNA = "always")

table(seer0to19$mcont66, seer0to19$In_Medicaid_Flag, useNA = "always")

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
```

@@

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19 %>% filter(Agerecodewithsingleages_and_100 !=18 ))

table1(~mcont66|age_cat, droplevels = TRUE, seer0to19%>% filter(Agerecodewithsingleages_and_100 <18 ))
```

@@

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()

# exclude those with missing stage and redefine some variables so the function doesn't throw an error.
seer0to19_stage <- seer0to19_stage %>%
  filter(!is.na(stage)) %>% # this yields 21,502
  select(stage, In_Medicaid_Flag, SEER_registry, mcont66, race_cat2, age_cat, Yost_ACS_2006_2010_quintile, state, can_type) %>%
  mutate(stage2 = case_when(stage == "In situ/localized" ~ "0: In situ/localized",
                            stage == "Regional" ~ "1: Regional",
                            stage == "Distant" ~  "2: Distant"))

@@

# relabel columns and dropt intercept rows
names(df4)[1:3]<-c("OR", "lowCI", "highCI")
df4

df4<-df4[-c(1,3),]
df4 <- tibble::rownames_to_column(df4, "Category") # Apply rownames_to_column

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
df4$Stage<- str_replace_all(df4$Category, rep_str)
df4$Category<- str_replace_all(df4$Category, rep_str2) # this is the dataframe for summary results

@@

df8<-df8[-c(1,5),]
df8 <- tibble::rownames_to_column(df8, "Category") # Apply rownames_to_column
df8

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
df8$Stage<- str_replace_all(df8$Category, rep_str)
df8
rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df8$Category<- str_replace_all(df8$Category, rep_str2)
df8 # this is the dataframe for summary results
```

# Adjusted models
```{r}
model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile , data=seer0to19_stage)
summary(model1a)
nrow(fitted(model1a)) # 21,507 is number of obs

@@

df12<-df12[-c(1,3:11, 13:20),]
df12 <- tibble::rownames_to_column(df12, "Category") # Apply rownames_to_column
df12

@@

# adjusted model
model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

@@

df16<-df16[-c(1,5:13, 17:24),]
df16 <- tibble::rownames_to_column(df16, "Category") # Apply rownames_to_column
df16

@@

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```

@@

# adjusted model
model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

@@

df20<-df20[-c(1,5:13, 17:24),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20  
```

@@

names(df2a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df2a<-df2a[-c(1,3:11, 13:20),]
df2a <- tibble::rownames_to_column(df2a, "Category") # Apply rownames_to_column
df2a  
df2a$CancerType<-i # add cancer type
df1a <-rbind(df2a, df1a)
}

@@

names(df4a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df4a <-df4a[-c(1,5:13, 17:24),]
df4a <- tibble::rownames_to_column(df4a, "Category") # Apply rownames_to_column
df4a  
df4a$CancerType<-i # add cancer type
df3a <-rbind(df4a, df3a)
}

@@

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant', '1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
SuppTable1$Stage<- str_replace_all(SuppTable1$Category, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
SuppTable1$Category<- str_replace_all(SuppTable1$Category, rep_str2)

SuppTable1
```

@@

@@

# FIGURE 2: Stage analysis by cancer type
```{r}
# graph ORs and 95% CIs
SuppTable1a <- SuppTable1 %>%
  filter(CancerType != "V. Retinoblastoma")

@@

SuppTable1a$Stage <- factor(SuppTable1a$Stage, levels = c("Regional", "Distant"))

@@

Plot1 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Regional") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A", x = "Cancer type") +
  ylim(0,2) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

@@

Plot2 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Distant") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B", x = "Cancer type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

@@

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1a$Category <- factor(SuppTable1a$Category, levels = c("Enrolled", "Discontinuous", "Other", "Continuous"))

>>>>

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$race_cat2<-factor(seer0to39$race_cat2, levels=c(0:2), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Other"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

@@

seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("in situ/localized", "Regional", "Distant"))

@@

@@

# Exclude those diagnosed before 2006 
```{r}
seer0to19 <- seer0to19 %>%
  filter(Year_of_diagnosis >=2006) # 32,380
```

@@

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) %>% # Other
  mutate(mcont66b = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb2 == 0 ~ 1, # continuous
                             mcontb2 == 1 ~ 2, # discontinuous
                             mcontb2 == 2 ~ 3, # Other
                             mcontb2 == 3 ~ 4)) # at or after diagnosis

table(seer0to19$mcont66, useNA = "always")
table(seer0to19$mcont66, seer0to19$In_Medicaid_Flag, useNA = "always")

table(seer0to19$mcont66b, useNA = "always")
table(seer0to19$mcont66b, seer0to19$In_Medicaid_Flag, useNA = "always")

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
seer0to19$mcont66b <- factor(seer0to19$mcont66b, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))

table(seer0to19$mcont66b)
```

@@

table1(~can_type|In_Medicaid_Flag, seer0to19)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
```

# FOR REVISION (insurance distribution)
```{r}
# classify primary payor variable
class(seer0to19$Primary_Payer_at_DX)
seer0to19$Primary_Payer_at_DX <- as.numeric(seer0to19$Primary_Payer_at_DX)

seer0to19 <- seer0to19 %>%
  mutate(payer = case_when(Primary_Payer_at_DX %in% c(1,2) ~ 0,
                           Primary_Payer_at_DX %in% c(10,20,21) ~ 1,
                           Primary_Payer_at_DX %in% c(31,35) ~ 2,
                           Primary_Payer_at_DX %in% c(60,64) ~ 3,
                           Primary_Payer_at_DX %in% c(65,66,67,68) ~ 4,
                           Primary_Payer_at_DX == 99 ~ 5))

seer0to19$payer <- factor(seer0to19$payer, levels = c(0:5), labels = c("Uninsured", "Private and Insurance NOS", "Medicaid", "Medicare", "Other", "Unknown"))

seer0to19$Insurance_Recode_2007 <- as.numeric(seer0to19$Insurance_Recode_2007)
table(seer0to19$Insurance_Recode_2007, useNA = "always")
seer0to19 <- seer0to19 %>%
  mutate(insurance = case_when(Insurance_Recode_2007 == 1 ~ 0,
                               Insurance_Recode_2007 == 2 ~ 1,
                               Insurance_Recode_2007 == 3 ~ 2,
                               Insurance_Recode_2007 == 4 ~ 3,
                               Insurance_Recode_2007 == 5 ~ 4))

label(seer0to19$insurance) <- "Insurance at diagnosis or initial treatment"

seer0to19$insurance <- factor(seer0to19$insurance, levels = c(0:4), labels = c("Uninsured", "Any Medicaid", "Insured", "Insured/No specifics", "Insurance status unknown"))

# Not shown in paper
table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2 + insurance|mcont66, droplevels = TRUE, seer0to19)

table1(~insurance|mcont66, droplevels = TRUE, seer0to19)
```

@@

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()
```

# FOR REVISION QUESTION (why not use AJCC staging vs. summary stage)--check missing data by cancer type for both
```{r}
colnames(seer0to19_stage)

table1(~ as.factor(SEERCombinedSummaryStage2000200) + as.factor(DerivedAJCCStageGroup6thed20042), seer0to19_stage)
```

```{r}
# exclude those with missing stage and redefine some variables so the function doesn't throw an error.
seer0to19_stage <- seer0to19_stage %>%
  filter(!is.na(stage)) %>% # this yields 21,502
  select(stage, In_Medicaid_Flag, SEER_registry, mcont66, race_cat2, age_cat, Yost_ACS_2006_2010_quintile, state, can_type) %>%
  mutate(stage2 = case_when(stage == "in situ/localized" ~ "0: In situ/localized",
                            stage == "Regional" ~ "1: Regional",
                            stage == "Distant" ~  "2: Distant"))

@@

# relabel columns and drop intercept rows
names(df4)[1:3]<-c("OR", "lowCI", "highCI")
df4

df4<-df4[-c(1,3),]
df4 <- tibble::rownames_to_column(df4, "Exposure comparison") # Apply rownames_to_column

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional vs. in situ/localized','2: Distant:In_Medicaid_FlagEnrolled'='Distant vs. in situ/localized')
rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
df4$Stage<- str_replace_all(df4$`Exposure comparison`, rep_str)
df4$`Exposure comparison`<- str_replace_all(df4$`Exposure comparison`, rep_str2) # this is the dataframe for summary results

@@

df8<-df8[-c(1,5),]
df8 <- tibble::rownames_to_column(df8, "Exposure comparison") # Apply rownames_to_column
df8

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional vs. in situ/localized','1: Regional:mcont66Discontinuous'='Regional. vs. in situ/localized','1: Regional:mcont66Other'='Regional vs. in situ/localized', '2: Distant:mcont66Continuous'='Distant vs. in situ/localized','2: Distant:mcont66Discontinuous'='Distant vs. in situ/localized','2: Distant:mcont66Other'='Distant vs. in situ/localized')
df8$Stage<- str_replace_all(df8$`Exposure comparison`, rep_str)
df8

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous vs. Not enrolled','1: Regional:mcont66Discontinuous'='Discontinuous vs. Not enrolled','1: Regional:mcont66Other'='Other vs. Not enrolled', '2: Distant:mcont66Continuous'='Continuous vs. Not enrolled','2: Distant:mcont66Discontinuous'='Discontinuous vs. Not enrolled','2: Distant:mcont66Other'='Other vs. Not enrolled')

df8$`Exposure comparison`<- str_replace_all(df8$`Exposure comparison`, rep_str2)
df8 # this is the dataframe for summary results
```

# Adjusted models
```{r}
model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
summary(model1a)
nrow(fitted(model1a)) # 21,502 is number of obs

@@

df12<-df12[-c(1, 3:23, 25:44),]
df12 <- tibble::rownames_to_column(df12, "Category") # Apply rownames_to_column
df12

@@

# adjusted model
model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

@@

df16<-df16[-c(1, 5:25, 29:48),]
df16 <- tibble::rownames_to_column(df16, "Category") # Apply rownames_to_column
df16

@@

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')

df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```

@@

# adjusted model
model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

@@

df20<-df20[-c(1, 5:25, 29:48),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20  
```

@@

names(df2a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")

df2a<-df2a[-c(1, 3:12, 14:22),]
df2a <- tibble::rownames_to_column(df2a, "Category comparison") # Apply rownames_to_column
df2a  
df2a$CancerType<-i # add cancer type
df1a <-rbind(df2a, df1a)
}

@@

names(df4a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")

df4a <-df4a[-c(1, 5:14, 18:26),]
df4a <- tibble::rownames_to_column(df4a, "Category comparison") # Apply rownames_to_column
df4a  
df4a$CancerType<-i # add cancer type
df3a <-rbind(df4a, df3a)
}

@@

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional vs. in situ/localized','1: Regional:mcont66Discontinuous'='Regional vs. in situ/localized','1: Regional:mcont66Other'='Regional vs. in situ/localized', '2: Distant:mcont66Continuous'='Distant vs. in situ/localized','2: Distant:mcont66Discontinuous'='Distant vs. in situ/localized','2: Distant:mcont66Other'='Distant vs. in situ/localized', '1: Regional:In_Medicaid_FlagEnrolled'='Regional vs. in situ/localized','2: Distant:In_Medicaid_FlagEnrolled'='Distant vs. in situ/localized')
SuppTable1$`Stage outcome comparison`<- str_replace_all(SuppTable1$`Category comparison`, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous vs. Not enrolled','1: Regional:mcont66Discontinuous'='Discontinuous vs. Not enrolled','1: Regional:mcont66Other'='Other vs. Not enrolled', '2: Distant:mcont66Continuous'='Continuous vs. Not enrolled','2: Distant:mcont66Discontinuous'='Discontinuous vs. Not enrolled','2: Distant:mcont66Other'='Other vs. Not enrolled', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
SuppTable1$`Category comparison`<- str_replace_all(SuppTable1$`Category comparison`, rep_str2)

names(SuppTable1)[1]<-c("Exposure comparison")

@@

# Replace multiple strings at a time
rep_str = c('Regional'='Regional vs. in situ/localized','Distant'='Distant vs. in situ/localized')
Table2$Stage<- str_replace_all(Table2$Stage, rep_str)

rep_str2 = c('Enrolled'='Enrolled vs. Not enrolled','Continuous'='Continuous vs. Not enrolled', 'Discontinuous'='Discontinuous vs. Not enrolled', 'Other'='Other vs. Not enrolled')
Table2$Category<- str_replace_all(Table2$Category, rep_str2)

names(Table2)[1]<-c("Exposure comparison")
names(Table2)[5]<-c("Stage outcome comparison")

# Reorder columns using the pipe operator
Table2 <- Table2 %>% select("Exposure comparison" ,  "OR", "cluster.lowCI","cluster.highCI", "CancerType", "Stage outcome comparison")

@@

# FIGURE 2: Stage analysis by cancer type
```{r}
# graph ORs and 95% CIs
SuppTable1a <- SuppTable1 %>%
  filter(`CancerType` != "V. Retinoblastoma")

@@

SuppTable1a$`Stage outcome comparison` <- factor(SuppTable1a$`Stage outcome comparison`, levels = c("Regional vs. in situ/localized", "Distant vs. in situ/localized"))

@@

Plot1 <- SuppTable1a %>%
  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1,  color = "black") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A. Regional vs. in situ/localized", x = "Cancer type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

@@

Plot2 <- SuppTable1a %>%
  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Distant vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1, color = "black") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B. Distant vs. in situ/localized", x = "Cancer type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

@@

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1a$`Exposure comparison` <- factor(SuppTable1a$`Exposure comparison`, levels = c("Enrolled vs. Not enrolled", "Discontinuous vs. Not enrolled", "Other vs. Not enrolled", "Continuous vs. Not enrolled"))

Plot3 <- SuppTable1a %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "C. Regional vs. in situ/localized", x = "Cancer type") +
  ylim(0,4.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                        values=c("red", "cyan3", "blue"))

Plot3
table(SuppTable1a$`Exposure comparison`)
